---
title: "FARS functions package Vignette"
author: "Federico Calore"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    fig_caption: yes
vignette: >
  %\VignetteIndexEntry{FARS functions package Vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This package provides some convenient functions to extract, aggregate and
visualise data from FARS - *Fatality Analysis Reporting System*.
More information can be found
[at this link](https://www.nhtsa.gov/research-data/fatality-analysis-reporting-system-fars).

```{r, include=FALSE}
library(buildRpackage)
datafile <- system.file("extdata", "accident_2013.csv.bz2", package = "buildRpackage")
```

## Low level data functions

The package includes some basic low-level functions to build a properly
formatted filename and read a csv file.
The data looks like the following:

```{r}
tabl <- fars_read(datafile)
tabl
```

## Data aggregations functions

Some convenient higher level functions are also available. These take as argument
a vector of integers to determine the years. If a corresponding year is found in
the data, the observations are aggregated by month and by year, and visualised
as a contigency table, with Months in rows and Years in columns, as the following:

```{r}
aggregated <- fars_summarize_years(c(2013, 2014, 2015), internal = TRUE)
aggregated
```

## Plot functions

Finally, a plot function is also available. It takes two integer arguments, one
for a USA State number and a second for the year.
The resulting observations are then plotted on the selected State map, using the
longitude and latitude information from the data file.
Here's a couple of examples:

```{r, fig.width=3, fig.height=3, ig.align='center', fig.show='hold'}
fars_map_state(1, 2013, internal = TRUE)
fars_map_state(48, 2015, internal = TRUE)
```
